DB pro sem - isolation levels

### 2020-12-13 paper.tex
updated paper

\subsection{Read uncommitted}
We define read uncommitted to be the isolation level that prohibits the phenomenon
P0, dirty write, this can be accomplished by placing long-duration write locks when modifying a data item.
Thus only allowing one transaction at a time to modify a given value.

@@

\includegraphics[width=8cm]{P5}
\begin{example}
    (first commit wins): \\
    Alice reads x and y to be 50€. Bob then reads the same data, while getting his own snapshot
    to operate on (newer timestamp than Alice). Alice then sets y to be -40€ and commits,
    resulting in a sum of 10€. Bob then sets y to be 30€ and tries to commit. As Alice already
    modified x Bob’s transaction will abort.
\end{example}

\subsection{Comparison with repeatable read}
In comparison to snapshot isolation repeatable read prohibits P5B while snapshot isolation prohibits some cases of
P3.
\subsection{Strength}
Snapshot isolation is comparable in strength to repeatable read as both prohibit P0, P1, and P2.
\subsection{Comparison to ANSI}
When looking at the weak interpretation of the ANSI specification as in \cite{Adya_Liskov_O_Neil_2000}
one can see that Snapshot isolation prohibits all three anomlies A1, A2, and A3 making it ANOMALY SERIALIZABLE.
To show this we will now take a look at A3 as we have already shown above that Snapshot isolation prohibits P1 and P2.
\subsection{A3 - Anomaly Phantom}
This is the strict interpretation of the Phatom anomaly. The difference to P3 is the position of the
transactions comit is not indiferent.
\begin{example}
    Alice reads the predicate blue cars. Bob then adds a new blue car to the database and increases
    the count of blue cars. Bob commits.  Alice then reads the count of blue cars. This leads to a phantom item
    (Alice can’t see Bob's car but knows of its existence due to the increased count).
\end{example}
Thus A3 only opperates on commited data while P3 accounts for transactions to still be in prosses.
\section{Conclusion}
\includegraphics[width=8cm, height=8cm]{iso_lvl_dia.png}

In conclusion one can see that the ANSI specification is inherintly flawed whwn talking
about multi-version histories and non locking database systems, as it failed to consider
these systems when defining the phenomena.
\bibliographystyle{plainnat}
\bibliography{refs}

>>>>

\subsection{Read uncommitted}
We define read uncommitted to be the isolation level that prohibits the phenomenon
P0, dirty write, this can be accomplished by placing long-duration write locks when modifying a data item.
This only allows one transaction at a time to write to a given value and requires all others to wait for a commit or abort.

@@

\includegraphics[width=8cm]{P5}
\begin{example}
    (first commit wins):
    Constrain: $ X+Y \geq  50$
    Alice reads x to be 10€ and y to be 50€. Bob then reads the same data, while getting his own snapshot
    to operate on (newer timestamp than Alice). Alice then sets x to be 5€,
    resulting in a sum of 55€. Bob then sets y to be 40€. Allice commits and Bob then tries to commit.
\end{example}
This would be prohibited by repeatable read but is allowed in snapshot isolation.
\subsection{Implications}
When looking at the phenomena defined above one can se that prohibiting it with a locked based system
would be the same as prohibiting P3 as a lost update can onl be prohibited by placing long locks.
But when using a multi version system instead proscribing P5B isn't equal to proscribing P3.

The example in 5.1 is allowed by Snapshot isolation as it does not check for constrains and the transaction
doesn't modify the same item (x and y respectively). Repeatable read on the other hand would disallow such
behaviour as it places long read and write locks on the items.
P5B is thus usefull to differntiate between Snapshot isolation and repeatable read.
\subsection{Comparison with repeatable read}
As staded above repeatable read prohibits P5B while snapshot isolation prohibits some cases of
P3 and both prohibit P0, P1, and P2. Although the argument above is made with a locking implementation
it is general as the phenomena defined were strengthened to match the locking behaviour.
\subsection{Strength}
Snapshot isolation is comparable in strength to repeatable read as both prohibit P0, P1, and P2.
To further investigate their difference we have to take a look at the original ANSI specification of serializable,
as it prohibits A3, which will be defined in the following section.
\subsection{A3 - Anomaly Phantom}
This is the strict interpretation of the Phatom anomaly. The difference to P3 is the position of the
transactions commit is not indiferent for A3.
\begin{example}
    Alice reads the predicate blue cars. Bob then adds a new blue car to the database and increases
    the count of blue cars. Bob commits.  Alice then reads the count of blue cars. This leads to a phantom item
    (Alice can’t see Bob's car but knows of its existence due to the increased count).
\end{example}
Thus A3 only opperates on commited data while P3 accounts for transactions to still be in prosses,
making A3 a weaker phenomenon than P3.
\subsection{Comparison to ANSI}
When looking at the weak interpretation of the ANSI specification as in \cite{Adya_Liskov_O_Neil_2000}
one can see that Snapshot isolation prohibits all three anomlies A1, A2, and A3 making it ANOMALY SERIALIZABLE.
To show this we will now take a look at A3 as we have already shown above that Snapshot isolation prohibits P1 and P2.
As the definition of A3 states the importance of commit sequence it is easy to see that Snapshot isolation with a first commit winns
stategy precluders A3, as a already altered item will lead to an abort.
When looking at a first write wins the same is true as still only one modified value will
be written to the database and the other one will be discarded by an abort. It thus only changes
which transaction winns in comparison to first commit wins.
\subsection{Comparison to Serializable}
In comparison to the serializable isolation level defined in 3.4 we can say that Snapshot isiolation is
weaker as serializable precluders P5B as well as P3, which as shown in example 5.1 above doesn"t hold true
for Snapshot isiolation.
\section{Conclusion}
\includegraphics[width=8cm, height=8cm]{iso_lvl_dia.png}

In conclusion one can see that the ANSI specification is inherintly flawed when talking
about multi-version histories and non locking database systems, as it failed to consider
these systems when defining the phenomena. As this will inevitably lead to different interpretations
and implementations it might be advisable to update the ANSI standard
according to the definitions given in \cite{Adya_Liskov_O_Neil_2000} and \cite{Berenson_Bernstein_Gray_Melton_O_Neil_O_Neil_1995}
to ensure the correct behaviour of non locking implementations of ANSI SQL.
\bibliographystyle{plainnat}
\bibliography{refs}
